\documentclass[x11names,reqno,14pt]{extarticle}
\input{preamble}
\DeclareMathOperator{\cl}{cl}
\usepackage[document]{ragged2e}
\usepackage{amsmath}
\pagestyle{fancy}{
	\fancyhead[L]{Winter 2023}
	\fancyhead[C]{221B - Cell Complexes}
	\fancyhead[R]{John White}
  
  \fancyfoot[R]{\footnotesize Page \thepage \ of \pageref{LastPage}}
	\fancyfoot[C]{}
	}
\fancypagestyle{firststyle}{
     \fancyhead[L]{}
     \fancyhead[R]{}
     \fancyhead[C]{}
     \renewcommand{\headrulewidth}{0pt}
	\fancyfoot[R]{\footnotesize Page \thepage \ of \pageref{LastPage}}
}
\DeclareMathOperator{\rel}{rel}

\title{221B - Cell Complexes}
\author{John White}
\date{Winter 2023}


\begin{document}

\section*{Lecture 1}

We will be using Hatcher's Algebraic Topology. The topology sequence is usually something like 
\begin{align*}
A & \text{Topological Spaces} \\
B & \text{Cell Complexes} \\
C & \text{Manifolds} \\
\end{align*}

\thm (BIG Theorem)

Given a ``reasonably nice" space, there is a bijection between connected covers of a space and subgroups of the fundamental group. 

\subsection*{\underline{Categories:}}

Algebraic structures that are much flabbier than a group. They consist of 
\begin{itemize}
\item A collection of arrows
\item A partial binary operation on these arrows
\item Objects, which arrows go between
\end{itemize}
We also want a composition law. That is, for objects and arrows
\[
\begin{tikzcd}
A \ar[r, "f"] & B \ar[r, "g"] & C 
\end{tikzcd}
\]
there is an arrow $\begin{tikzcd} A \ar[r, "g \circ f"]&  C \end{tikzcd}$. We want this composition to be associative, that is $(f \circ g) \circ h = f \circ (g \circ h)$, and we want objects to have identity arrows. 

Not all functions have inverses. Using sets and functions as an example, we have described the category Set. 

Here are some more examples of categories: 
\exm

\begin{itemize}
\item Groups and group homomorphisms (Grp)
\item Topological spaces and continuous functions (Top)
\item etc. 
\end{itemize}

We can make the following new category. 

\defn

We denote by $\Top^*$ the \underline{category of based topological spaces}, whose objects are pairs $(X, x_0)$, where $X$ is a topological space and $x_0 \in X$, and whose morphisms are continuous functions $f:(X,x_0)\to(Y, y_0)$ such that $f(x_0) = y_0$. 

\subsection*{Goal:}

Our goal is to get a functor from Top to Grp. The fundamental group functor $\pi_1$ will go from $\Top^*$ to Grp.

\section*{Lecture 2}

\subsection*{\underline{Topology review:}}

\defn 

A \underline{topological space} is a set $X$ along with a collection of subsets of $X$ called ``open sets," such that $X, \varnothing$ are open, and the arbitrary union and finite intersection of open sets are open. 

Notice the following diagram commutes using the product topology

\[
\begin{tikzcd}
&\ar[ld, "f"'] \ar[d, dotted, "\exists !"]Z \ar[rd, "g"] &  \\
X  &\ar[l, "P_X"'] X\times Y \ar[r, "P_Y"] & Y\\
\end{tikzcd}
\]

And in general 

\[
\begin{tikzcd}
Z \ar[d, dotted, "\exists !"']  \ar[rd, "f_\alpha"] & \\
\prod_{\alpha\in A}X_\alpha \ar[r, "P_\alpha"]& X_\alpha\\
\end{tikzcd}
\]
Maps are continuous; functions are not. 

\lem (Gluing lemma)

Suppose $f:A\to Y$, $g:B\to Y$ are continuous, and $f(x) = g(x)$ for all $x \in A \cap B$. Then $f\cup g: A \cup B \to Y$ is continuous. This only holds as long as $A, B \subseteq X$ are closed. 

\subsection*{\underline{Same Shape, Same Map}}
(maps up to wriggling things around a bit)

\defn Two maps are \underline{homotopic} if there exists a parametrized map $f_t:X\to Y$ such that $f_0 = f, f_1 = g$ for $f, g:X \to Y$. Equivalently, and more precisely, if there exists a map $F:X\times[0,1]\to Y$ such that $F(x, 0) = f(x), F(x, 1) = g(x)$ for all $x \in X$. 

$X, Y$ topological spaces are said to have the same shape if there exist maps $f:X\to Y, g:Y\to X$ such that $g\circ f\simeq \Id_X$ and $f\circ g \simeq \Id_Y$. We may say that $X, Y$ have the same \underline{homotopy type}

\defn

A \underline{deformation retraction} from $X \to A\subseteq X$ is a map from $X \times I \to X$ such that, for all $x \in A$, and $s, t \in I$,
\begin{align*}
f_0(x) &= x&\forall x \in X \\
f_1(x) &\in A&\forall x \in X \\
f_t(x) &= f_s(x)& \forall x \in A \\
\end{align*}

\section*{Lecture 3, 1/13/23}

\defn

Let $X$ be a topological space. A \underline{retraction} is a map $r:X\to X$ such that $r \circ r = r$. That is, $r(r(x)) = r(x)$ for any $x \in X$. 

Let $A = r(X)$. Then $r|_A = \Id_A$. 

\defn

Let $F:X\times I\to Y$. We say $f_0 \simeq f_1 \rel A\subseteq X$ are \underline{homotopic relative to $A$} if, for any $x \in A$, $f_t(x)$ is independent of $t$. That is, for any $s, t \in I$, $f_s(x) = f_t(x)$ for any $x \in A$. 


For any map $f:X\to Y$, there exists a space $Z \simeq Y$ via $g:Y \to Z$ such that $g \circ f:X\to Z$ is injective. That is, in the following diagram, we have a bijection between homotopy classes of maps $f$ and homotopy classes of maps $g\circ f$, and we can do this in a way that rigs $g \circ f$ to be injective. 
\[
\begin{tikzcd}
X \ar[r, "f"] \ar[rd, "g\circ f"'] & Y \ar[d, "g"] \\
& Z \\
\end{tikzcd}
\]

\defn

Given a map $f:X\to Y$ we can construct the \underline{Mapping Cylinder} $M_f$ by setting $M_f = X \times I \coprod Y/\sim$, where $(x, 0) \sim f(x)$. 

The visual intuition should be taking the disjoint union of $X$ and $Y$, and tieing a string between $x$ and $f(x)$ for each point. 

\claim 
$X \hookrightarrow M_f, Y \hookrightarrow M_f$, and the latter is in fact a homotopy equivalence. Further, the injection $X \hookrightarrow M_f$ is homotopic to $f(X) \hookrightarrow M_f$. 
\proof
You can construct a homotopy which ``squishes" the cylinder down to $f(X)$. 

\qed

\defn

A space $X$ is \underline{contractible} if it has the homotopy type of a point. A map is \underline{null-homotopic} if it is homotopic to a constant map. So $X$ is contractible if the identity is null-homotopic. 

Now he's drawing an example. The example is Bing's House with 2 rooms, which I will not reproduce here. But the point is that it's contractible, but not obviously so. 

\subsection*{\underline{Cell Complexes}}

Cell complexes are topological spaces which are built up inductively out of closed balls in Euclidean space. We write $\mathbb{D}^n := \{\vec{x} \in \R^n \mid \norm{\vec{x}} \leq 1 \}$, and $e^n := \{\vec{x}\in\R^n \mid \norm{\vec{x}} < 1\}$. We can see that $e^n = \operatorname{int}\mathbb{D}^n$, and $\del\mathbb{D}^n = \mathbb{S}^{n - 1}$. 

\subsubsection*{\underline{Base step}}

Start with some collection of points $X^0$, the $0$-skeleton, with the discrete topology.

\subsubsection*{\underline{Inductive step}}

Let $X^{n - 1}$ be the $n - 1$ skeleton, which has already been build and defined. Select some collection of $n$-dimensional balls $\{\mathbb{D}^n\}_{\alpha\in A}$, and some continuous ``attaching map" $\varphi_\alpha:\del\mathbb{D}^n_\alpha\to X^{n - 1}$. Then 
\[
(X^n = X^{n - 1}\coprod_{\alpha\in A}\mathbb{D}^n )/(x \sim\varphi_\alpha(x)\forall x \in \del\mathbb{D}^n)
\]

\section*{Lecture 4, 1/18/23}

A space $X$ is a \underline{cell complex} if it has been constructed using the above inductive procedure. If $n = \oo$, we use the weak topology, in which the open sets are the sets which are open when intersected with each $X^n$. 

For every $\mbb{D}^n_\alpha$ and corresponding ``attaching map $\varphi_\alpha:\del\mbb{D}^n_\alpha\to X^{n - 1}$, there is a subset of $X^n$ homeomorphic to $\operatorname{int}(\mbb{D}^n_\alpha)$, via the composition
\[
\operatorname{int}(\mbb{D}^n_\alpha) \hookrightarrow \mbb{D}^n_\alpha \hookrightarrow X^{n - 1}\coprod_{\alpha}\mbb{D}^n_\alpha \to X^n
\]
which we call $\Phi_\alpha:\mbb{D}^n_\alpha\to X^{n - 1}$. So the attaching map $\phi_\alpha:\del\mbb{D}^n_\alpha\to X^{n - 1}$ extends to a ``characteristic map" $\Phi_\alpha$.

We will now see many examples of things. 

\exm

If you stop after constructing $X^1$, it's a graph. 

\exm

$\mbb{S}^n$ has a cell structure with one $e_0$ and one $e_n$. 

\exm

Consider $\RP^2$. This can be expressed as $(\R^3\setminus\{0\})/(\vec{x}\sim\lambda\vec{x}, \lambda\neq0)$. We can replace $2$ with any $n$ and get $\RP^n$. Indeed, we can replace $\R$ with $\C$, $\mbb{H}$, or indeed any field. 

\underline{Homogenous coordinates}

For $(x, y, z)\neq(0,0,0)$, we have $[x, y, z]\eqdef\{(\lambda x, \lambda y, \lambda z) \mid \lambda \neq0\}$. For example, $[1,2,3]=[2,4,6]$. 

\section*{Lecture 5, 1/20/23}

\defn

A \underline{subcomplex} of a complex $X$ is a closed disjoint union of open cells $e^{n_i}_{\alpha_i}$ in $X$ such that they form a cell complex on their own. 

\rem We keep talking about ``CW Complexes." The C is for ``closure finite," and the W is for "weak topology." 

Recall: $\RP^n \eqdef \R^n/(x\sim\lambda x, \lambda\neq0)$. $\CP^n$ can be defined similarly. 

We can write $\RP^n = e^0 \cup e^1 \cup e^2 \cup \cdots \cup e^n$, and $\CP^n = e^0 \cup e^2 \cup e^3 \cup \cdots \cup e^{2n}$. We can do the same thing with the quaternions. 

Next time, we will cover operations on complexes. 

\section*{Lecture 6, 1/23/23}

This lecture, we will cover operations on cell complexes, and two big theorems.

\subsection*{\underline{Operations on Cell Complexes}}

\begin{enumerate}

\item If $X, Y$ have cell structures, then $X \times Y$ has a natural cell structure.

\item If $(X, A)$ is a CW-pair, then $X/A$ has a natural cell structure ($X/A$ denotes identifying all points in $A$ together). 
\[
\begin{tikzcd}
\mbb{D}'_\alpha \supseteq \del\mbb{D}'_\alpha \ar[d, "\phi_\alpha"] \ar[dr] & \\
X^0 \ar[r, "q"'] & (X/A)^0 \\
\end{tikzcd}
\] 

\item Cones and Suspensions. The cone on $X$, $CX$, is defined as
\[
CX = (X\times I)/(X\times\{0\})
\]
Note that $CX$ is contractible for any $X$. The suspension on $X$, $SX$, is defined as 
\[
SX = CX/(X\times\{1\})
\]

If $f:X\to Y$ is a map, there exists a natural map $Sf:SX\to SY$. Indeed, if $f:X\to Y$, then $f\times\Id:X\times I\to Y\times I$, and so we can factor $f \times \Id$ through the quotient map:
\[
\begin{tikzcd}
X\ar[r, "f"]\ar[d, hook]\ar[dr, dotted] & Y\ar[d, hook] \\
X\times I\ar[d,"q"]\ar[r, "f\times\Id"]\ar[dr, dotted] & Y\times I \ar[d] \\
SX\ar[r, "\exists!Sf"'] & SY \\
\end{tikzcd}
\]
Note $S(\mbb{S}^n) = \mbb{S}^{n + 1}$.

\item Joins. If $X, Y$ are cell structures, then we define their join $X\star Y$ as
\[
X\star Y = \frac{X\times Y \times I}{(x, y_1, 0) \sim (x, y_2, 0), (x_1, y, 1) \sim (x_2, y, 1)}
\]
This is a useful construction for simplices. 

\item Wedge product. If $X, Y$ are cell structures, with distinguished points $x_0 \in X, y_0 \in Y$, then we define their wedge product $X \wedge Y$ as
\[
X\wedge Y = \frac{X\coprod Y}{x_0 \sim y_0}
\]
This is just gluing $X$ and $Y$ together at a distinguished point. This raises an obvious question: does the wedge product depend on the points $x_0, y_0$? Yes, but not if they are (connected) cell complexes!

If $x_0$ is a 0-cell of $X$, and $y_0$ a 0-cell of $Y$, then $X \vee Y$ has a natural cell structure AND $X\vee Y$ is a subcomplex of $X\times Y$.

\item Smash product. If $X, Y$ are spaces with distinguished points $x_0, y_0$, then the smash product is defined as 
\[
X\wedge Y = \frac{X\times Y}{X\vee Y}
\]
For example, the smash product $S^1\wedge S^1$ is a Torus quotiented out by the longitudinal and meridian circles. By arguing from some cell nonsense, we can say this is $S^2$. 

\end{enumerate}

Here are two big theorems. 

\thm If $(X, A)$ is a CW-pair, and $A$ is contractible, then $X/A$ is homotopy equivalent to $X$, with the quotient mapping itself providing a homotopy equivalence. 

\thm Suppose $(X_1, A)$ is a CW-pair, and $f, g:A\to Y$ are maps. If $f \simeq g$, and everything in sight is a cell complex, then
\[
X_1\coprod_f Y \simeq X_1 \coprod_g Y
\]
That is, if $f, g$ are used as attaching maps, then the resulting spaces will be homotopy equivalent. 

\section*{Lecture 7, 1/25/23}

\defn

Let $X$ be a cell complex. If we let $f_i$ be the number of $i$-dimensional cells in the cell structure, then we define
\[
\chi(X) = f_0 - f_1 + f_2 - f_3 + \cdots
\]
The more general definition is the alternating sum of the Betti numbers of $X$, where the $i$th Betti number is $\dim H^i(X)$. 

\defn

Let $X$ be a topological space, and let $A\subseteq X$ be a subspace. We say that $(X, A)$ has the \underline{homotopy extension property} (HEP) if for all topological spaces $Y$ and for all maps $f:X\times\{0\} \cup A \times I \to Y$, there exists an extension of $f$, $\bar{f}:X\times I \to Y$, such that $\bar{f}|_{X\times\{0\}\cup A\times I} = f$. 

Slogan: ``A homotopy on the subspace can be extended to a homotopy on the entire space." 

\section*{Lecture 8, 1/27/23}

\prop

$(X, A)$ has the homotopy extension property if and only if $X \times I$ retracts to $X\times\{0\} \cup A \times I$.

\proof

\qed

\exm 

Does $(\mbb{D}^2, \del\mbb{D}^2)$ have the property? Does $\mbb{D}^2\times I$ retract onto $\mbb{D}^2\times\{0\}\cup \del\mbb{D}^2 \times I$? Yes. This is easy to see by drawing a picture. 

Here is a non-example. Let $X = I$, and let $A = \{\frac{1}{n}\}_{n\in\N}\cup\{0\}$. $X \times I$ is the square, and $X \times\{0\}$ is the bottom of the square, so $X\times \{0\} \cup A\times I$ is the comb space. The square doesn't retract to this. 

\prop 

If $(X, A)$ is a CW pair, then $(X, A)$ has the homotopy extension property. 

\proof

Later

\qed

\thm 

If $(X, A)$ has the homotopy extension property, and $A$ is contractible, then the quotient map $X \to X/A$ is a homotopy equivalence. 

\proof

Consider identity map $\Id:A\to A$. We have a homotopy $F:A\times I \to A$ which is a witness to $A$ being contractible. That is, $f_0 = \Id_A$, $f_1 \equiv \{p\}$ for some point $p \in A$. 

Then there is an extension to a homotopy $H:X\times I \to X$. We have the following commutative diagram
\[
\begin{tikzcd}
X\ar[r, "f_t"] \ar[d, "q"] & X \ar[d, "q"] \\
X/A \ar[r, "\bar{f_t}"] & X/A \\
\end{tikzcd}
\]
Because all of $A$ goes to a point for $t = 1$, then by the universal property of quotients, there is a map $g$ making the diagram commute:

\[
\begin{tikzcd}
X\ar[r, "f_1"] \ar[d, "q"] & X \ar[d, "q"] \\
X/A \ar[ur, "g"]\ar[r, "\bar{f_1}"'] & X/A \\
\end{tikzcd}
\]
So $qg$ is homotopic to the identity map, and $gq$ is homotopic to the identity map. This completes the proof? 

\qed

\section*{Lecture 9, 1/30/23}

\defn

We say that $(X, A)$ and $(Y, A)$ are homotopy equivalent relative to $A$ if there exist maps $f:X\to Y, g:Y\to X$ such that $f|_A = \Id_A, g|_A = \Id_A$ and $g \circ f \simeq \Id_X$ relative to $A$, and $f \circ g \simeq \Id_Y$ relative to $A$. 

\thm

If $(X, A)$ is a CW Pair, and $f, g:A\to X_0$ are homotopic maps, then $X_0 \coprod_fX_1 \simeq X_0\coprod_gX_1$ relative to $X_0$. 

\proof

Bunch of pictures I can't write down. 

\qed

\prop

If $(X, A), (Y, A)$ both have the homotopy extension property, and $f:X\to Y$ is a homotopy equivalence such that $f|_A = \Id_A$, then $f$ is a homotopy equivalence relative to $A$. 

\proof

\qed

\cor

If $(X, A)$ has the homotopy extension property and $A \hookrightarrow X$ is a homotopy equivalence, then $X$ deformation retracts to $A$. 

\proof

\qed

\cor

A map $f:X\to Y$ is a homotopy equivalence if and only if $X$ is a deformation retraction of $M_f$. 

\proof

\qed

\section*{Lecture 10, 2/3/23}

\defn 

Given any path $f:I\to X$, we write $[f]$ for the set of paths $g:I\to X$ such that $g \simeq f$ relative to $\del I$. If we don't fix endpoints, each $[f]$ would be a path-component. Sometimes we use $\pi^0$ to denote the set of path components. 

Let $f, g:I\to\R^n, f\simeq g$ by $h_t(u) = tg(u) + (1 - t)f(u)$. Then $h_0 = f, h)1 = g$. $h$ is called the ``straight line homotopy."

In fact, we could change $I$ to \underline{any} topological space, and $f \simeq g$ \underline{still} 

We could also change $\R^n$ to any $U\subseteq\R^n, U$ convex, and $f\simeq g$ \underline{still}. 

We could change $X$ to $U$, any metric space iwth unique shortest path which vary continuosly as the endpoints vary. 

Concatenation: If $f(1) = g(0)$, then define $f\star g:I\to X$ by 
\[
(f\star g)(t) = \begin{cases} f(2t) & t\in[0,\frac{1}{2}] \\ g(2t-1) & t \in [\frac{1}{2}, 1] \end{cases} 
\]

\defn

Assume $f(1) = g(0)$. Then $[f]\star[g] = [f\star g]$ is well defined by handwaving. 

Constants: The constant path $c_x$ is the path which is constantly $x$. Note $[c_{f(0)}]\star[f] = [f], [f]\star[c_{f(1)}] = [f]$. 

Inverses: Define $\bar{f}(u) = f(1 - u)$. Note $[f][\bar{f}] = [c_x]$. 

Associativity: $(f\star g)\star h \simeq f\star(g\star h)$. 

\defn A category where every $f:A\to B$ has an inverse (i.e. an arrow $f^{-1}:B\to A$ such that $ff^{-1} = \Id_B = f^{-1}f)$ is called a \underline{groupoid}. 

$\pi_0$ is a functor, (objects$\to$objects, arrow$\to$arrows, compositions$\to$compositions, identities$\to$identities)

$(X, x_0) \mapsto \pi_1(X, x_0) = \{[f]\mid f:I\to X, f(0) = f(1) = x_0\}$. 

$\begin{tikzcd} (X, x_0)\ar[r, "\pi_1"]&\pi_1(X, x_0) \end{tikzcd}$. 
\[
\begin{tikzcd}
(X, x_0)\ar[d, "h"]\ar[r, "\pi_1"]&\pi_1(X, x_0) & f \ar[d] \\
(Y, y_0) \ar[r, "\pi_1"]&\pi_1(Y, y_0) & h\circ f \\
\end{tikzcd}
\]

\section*{Lecture 11, 2/6/23}

just homework review

\section*{Lecture 12, 2/8/23}

Review for the midterm

Types of questions:

\begin{enumerate}

\item State Definitions (particularly important and complicated definitions)

\item Carefully state important theorems

\item Give key counterexamples

\item Prove easy propositions

\item Do simple constructions/modifications (applying our theorems)

\end{enumerate}

HEP: For any $f:X\times\{0\}\cup A\times I$, there exists an extesnsion $\bar{f}:X\times I \to Y$. This is equivalent to $X\times I$ def retracting to $X\times\{0\}\cup A\times I$
\[
\begin{tikzcd}
X\times\{0\}\cup A\times I \ar[d, "\iota"] \ar[r, "f"] & Y \\
X\times I \ar[u, "r"] \\ & 
\end{tikzcd}
\]

Questions from previous exams;

1. Carefully define the notion of a cell complex. 

2. Let $f:S^1\to S^1$ be continuous. Define the mapping cylinder $M_f$ and describe an explicit cell structure on it. 

3. What does it mean to say that computing the fundamental gorup is a functor from $\Top*$ to $\Grp$

4. In each case, find a simpler space with the same homotopy type, briefly explain reasons and use pictures. 
	a. suspension of disjoint union of 3 circles as a wedge product. 

	b. View $S^2$ as a subspace of $S^3$ and describe the quotient $S^3/S^2$. 

	c. Remove both the Z axis and the unit circle in the xy plane and describe a 2-dimensional object with the same homotopy type

5. Let $X$ be a topological space, prove that $f:S^1\to X$ extends to a map $F:D^2\to X$ if and only if $f$ is nullhomotopic. 

6. Define the homotopy extension property and then prove that a pair $(X, A)$ has the property if and only if there is a retraction $X\times\{0\}\cup A\times I$. 

7. Give examples of each of the following:
		retract of a cell complex onto a cell complex which does not extend to a def retraction

		Give an example of a contractible space that does not deformation retract to a point. 

\section*{Lecture 13, 2/13/23}

\prop

Let $h:I\to X$ be a path from $x_1$ to $x_0$. The map $\beta_h:\pi_1(X, x_1) \to \pi_1(X, x_0)$ defined by $\beta_h([f]) = [hf\bar{h}]$ is an isomorphism. 

\proof

\qed

\rem As long as $X$ is path-connected, $\pi_1(X, x_0)$ is well-defined, up to isomorphism, independent of $x_0$, but this isomorhpism is not canonical. 

If $x_1 = x_0$, then $h$ is a loop, and $[h]\in\pi_1(X, x_0)$. 

So $[f]\mapsto[hf\bar{h}] = [h][f][h]^{-1}$. So $\beta_h$ is an inner automorphism of $G = \pi_1(X, x_0)$. 

\defn

A map $p:\hat{X}\to X$ is called a \underline{covering map} if there exists an open cover $\{U_\alpha\}$ of $X$ such that for all $\alpha$, $p^{-1}(U_\alpha)$ can be decomposed into a disjoint union of subsets which are each homeomorphic to $U_\alpha$ and sent homeomorphically to $U_\alpha$ by $p$. 

Fact: If $p:\hat{X}\to X$ is a covering map and $p(\hat{x_0}) = x_0$, then the induced map $p_\star:\pi_1(\hat{X},\hat{x_0}) \to \pi_1(X, x_0)$ is injective. 

\section*{Lecture 14, 2/15/23}

\defn

If $X$ is a path connected space, then $X$ is said to be \underline{simply connected} if $\pi_1(X, x_0) = 0$. 

\prop

$X$ is simply connected if and only if the homotopy class of a loop is determined entirely by its endpoints. 

That is, if $\gamma, \sigma:[0,1]\to X$ are paths, and $\gamma(0) = \sigma(0), \gamma(1) = \sigma(1)$, then $\gamma \simeq \sigma$. 

\proof

\qed

Let $f, g$ be paths as in the statement of the problem, and let $h:[0,1]\to X$ be a path with $h(0) = f(1) = g(1), h(1) = f(0) = g(0)$. Because $\pi_1(X, x_0)$ is trivial, we know that $f\star h$ is contractible relative to $x_0$. The same is true of $h \star g$. So 
\begin{align*}
[f] & = [f\star e]\\
	 & = [f\star(h\star g)] \\
	 & = [(f \star h) \star g]\\
    & = [g] \\
\end{align*}

\qed

\defn

Given $p:Y\to X$ and $f:Z\to X$, a map $\tilde{f}:Z\to Y$ such that $p\circ\tilde{f} = f$, then $\tilde{f}$ is called a \underline{lift of $f$}.

\[
\begin{tikzcd}
Z\ar[dr, "f"]\ar[d, dotted, "\tilde{f}"'] & \\
Y \ar[r, "p"] & X \\
\end{tikzcd}
\]


\thm

Let $\Phi:\Z\to\pi_1(S^1, (1, 0))$ be given by $n\mapsto[\omega_n]$, where $\omega_n:I\to \mbb{S}^1$ is defined by $s\mapsto(\cos(2\pi ns), \sin(2\pi n s))$ (this is the path which wraps around $n$ times) is an isomorphism. 

\proof

One can think of the projection map $\rho:\R\to\mbb{S}^1$ given by $\rho(s) = e^{2\pi i s}$ as a projection from a ``spiral" above the circle. So we have a map 
\[
\begin{tikzcd}
\{0\} \ar[r]\ar[d] & \R \ar[d, "\rho"] \\
\,[0, 1]\ar[r, "f"]\ar[ur, dotted, "\exists! \tilde{f}"] & \mbb{S}^1 \\
\end{tikzcd}
\]

We need two facts: 

\begin{enumerate}[label=(\alph*)]

\item For all $f:I\to\mbb{S}^1$ starting at $x_0 \in \mbb{S}^1$, and for each choice of $\tilde{x_0} \in p^{-1}(x_0)$, there exists a unique path $\tilde{f}:I\to\R$ starting at $\tilde{x_0}$ which lifts $f$. 

\item For all $\underbrace{f_t:I\to\mbb{S}^1}_{F:I\times I\to\mbb{S}^1}$ starting at $x_0 \in \mbb{S}^1$, and for each choice of $\tilde{x_0} \in p^{-1}(x_0)$, there exists a unique lifted homotopy of paths $\tilde{f_t}:I\to \R$ which is a lift of $f_t$. 

\item For every $F:Y\times I\to \mbb{S}^1$ and every $\tilde{F}|_{Y\times\{0\}}:Y\times\{0\}\to\R$ which lifts $F|_{Y\times\{0\}}:Y\times\{0\}\to\mbb{S}^1$, there exists a unique $\tilde{F}:Y\times I\to\R$ lifting $F$. 

\[
\begin{tikzcd}
Y\times\{0\}\ar[d, hook]\ar[r, "\tilde{F}|_{Y\times\{0\}}"] & \R \ar[d, "p"] \\
Y\times I \ar[ur, dotted, "\exists!\tilde{F}"]\ar[r, "F"'] & \mbb{S}^1 \\
\end{tikzcd}
\]
\end{enumerate}

The first guarantees surjectivity of $\Phi;$ the second injectivity. 

Indeed, let $[f]$ be a homotopy class of maps based at $x_0$. Then, once we choose an element of $p^{-1}(x_0)$, this will lift to a path $[\tilde{f}]$ in $\R$ which is homotopic to one of the $\omega_n$'s. The second will then allow us to use that homotopy to extend to a homotopy between $f$ and $p(\omega_n)$.

Note: There exists a lift of $\omega_n$ starting at $(1, 0, 0)$ in the spiral, $\tilde{\omega_n}:I\to\R\subseteq\R^3, s\mapsto (\cos(2\pi n s), \sin(2\pi n s), ns)$. 

We will prove (c), which will prove the first two. 

There exists an open cover $\{U_\alpha\}$ of $\mbb{S}^1$ such that for all $\alpha, p^{-1}(U_\alpha) = \coprod_{\beta}\tilde{U}_\alpha^{(\beta)}$ where for each $\tilde{U}_\alpha^{(\beta)}$, 
\[
p|_{\tilde{U}_\alpha^{(\beta)}}: \tilde{U}_\alpha^{(\beta)} \to U_\alpha
\]
is a homeomorphism. 

\rem

This means that 
\[
(p|_{\tilde{U}_\alpha^{(\beta)}})^{-1}:U_\alpha\to\tilde{U}_\alpha^{(\beta)}
\]
exists and is continuous. 

From here we argue by compactness. 

\qed

\section*{Lecture 15, 2/17/23}

We know that $\pi_1$ is functor from the category of based topological spaces to the category of groups. This tells us that if $A \hookrightarrow X$ is a homotopy equivalence, i.e. if $X$ deformation retracts to $A$, with retraction $r$, then because $\iota\circ r = \Id_A$, we get that $\iota_*$ must be injective. 

So if $A \subseteq X$ admits a retraction, then $\pi_1(A, x_0) \subseteq \pi_1(X, x_0)$

\cor

If $X$ deformation retracts to $A$, then $\begin{tikzcd} \pi_1(A, a) \ar[r, "\iota_*"] & \pi_1(X, a)\end{tikzcd}$ is not just injective but is an isomorphism. 

\proof 

It is easy to see that any element of $\pi_1(X, a)$ homotopes to an element of $\pi_1(A, a)$. $\iota_*$ then clearly inverts this, so we have an isomorphism. 

\qed

We only know two facts:

\begin{enumerate}

\item $\pi_1(\mbb{S}^1) \cong \Z$ 

\item $\pi_1(\text{a single point}) = 0$

\end{enumerate}

Just this fact tells us that there isn't a retraction from the disk to its boundary, because then $\Z$ would inject into the trivial group, which is clearly not the case. 

This machinery finally will give us ways to prove negative statements - statements to the effect of ``there is no continuous $f:X\to Y$ with such and such property".


\thm (Fundamental theorem of algebra)

Let $p(z) = z^n + z_1z^{n - 1} + \cdots + a_n$, $p:\C\to \C$ has a root. 

\proof

Suppose $p$ has no roots. Fix a positive real $r$. Then define the map $\mbb{S}^1\to\mbb{S}^1$ by sending $s \in [0, 1]$ to $p(re^{2\pi i s})/p(r)$. This starts and ends at $1 \in \C$. We know this is not zero, so we can divide by the length to get 
\[
f_r(s) = \frac{p(re^{2\pi i s})/p(r)}{\norm{p(re^{2\pi i s})/p(r)}}
\]
As an element of $\pi_1(\mbb{S}^1, 1) \cong \Z$, if $r$ is small enough, then $[f_r(s)] = 0$. But if $r$ is large enough, then $[f_r(s)]= [z^n]$. This is because we can consider $p(z) = z^n + b(a_1z^{n - 1} + \cdots + a_n)$ and letting $b\to 0$. For this to work $r$ must be large enough.

But $F:[0,1]\times\mbb{S}^1\to\mbb{S}^1$ given by $F(r, s) = f_r(s)$ is a homotopy between $0$ and $[z^n]$, which is impossible. 

Thus $p$ must have a root. 

The punchline is essentially that as the radius grows large, it must run into a root eventually, or this bad behavior will happen.

\qed

\thm (Brouwer Fixed Point)

If $f:D^2\to D^2$ is continuous, then there exists $x \in D^2$ such that $f(x) = x$. 

\proof

Suppose this was not the case. Let $f$ be such that for all $x, f(x)\neq x$. Define the function $g:D^2\to\mbb{S}^1$ by letting $g(x)$ be the point on $\mbb{S}^1$ where the line going from $f(x)$ to $x$ intersects it. This is continuous and well-defined. This is also a retraction $r:D^2\to\mbb{S}^1$, which is impossible by previous discussion. 

\qed

\thm (Borsuk-Ulam)

For every function $f:\mbb{S}^2\to\R^2$, there is some $x \in \mbb{S}^2$ such that $f(x) = f(-x)$. In other words, if we flatten a sphere, there are a pair of antipodal points which get sent to the same point. 

\proof

Suppose that this is not the case. Let $f$ be a function $\mbb{S}^2\to\R^2$ such that $f(x)\neq f(-x)$ for all $x$. Define $g:\mbb{S}^2\to\mbb{S}^1$ by 
\[
x\mapsto \frac{f(x) - f(-x)}{|f(x) - f(-x)|}
\]
Note $g(-x) = -g(x)$. Consider the function $h:s\mapsto(\cos(2\pi s), \sin(2\pi s), 0)$, which is an equatorial loop around $\mbb{S}^2$, based at $(1, 0, 0)$. Then $gh:I\to S^1$ is a loop based at $(1, 0)$, and $h(s) = - h(s + \frac{1}{2})$. 

We now lift to the spiral $S$:

\[
\begin{tikzcd}
& S \ar[d]\\
I\ar[ur, "\tilde{h}"] \ar[r, "h"] & \mbb{S}^1\\
\end{tikzcd}
\]
We see that $\tilde{h}(s), \tilde{h}(s + \frac{1}{2})$ must differ by a half integer, i.e. $\tilde{h}(s) = \tilde{h}(s + \frac{1}{2}) + \frac{q}{2}$ with $q$ an odd integer (this can be seen by drawing the spiral and looking at lifts of antipodal points). 

So $\tilde{h}(1) = \tilde{h}(\frac{1}{2}) + \frac{q}{2}$. Because the integers are discrete, $q$ can't change as $s$ varies. Further, $\tilde{h}(0) = \tilde{h}(\frac{1}{2}) + \frac{q}{2}$, so $\tilde{h}(0) = \tilde{h}\frac{1}{2}) + \frac{q}{2} = \tilde{h}(1) + q$. So $[h] = -q$ is odd. So $[h]$ is not zero, so is a loop which is not contractible. 

Hitting everything with $\pi_1$, get a map from $\pi_1(I)$ to $\pi_1(\mbb{S}^1)$ which has image consisting of at least two points. But this is impossible, as $\pi_1(I) = 0$. So we have a contradiction, so we have finished the proof. 

\qed

\prop

For all $n \geq 2$, $\pi_1(\mbb{S}^n) = 0$. 

\proof

Let $\gamma:I\to\mbb{S}^2$ be a loop. If this loop misses one point, then taking away that point, then this is a loop in $\R^2$, which is contractible, so the loop is contractible. 
 
What if $\gamma$ is very poorly behaved? Well, then there is an open set whose preimage is a union of open intervals in $I$. Then some stuff happens idk. 

\qed

\thm (Invariance of domain)

$\R^2 \not\cong \R^m$ for $m > 2$. 

\proof

I missed it :(

\qed

\section*{Lecture 16, 2/22/23}

\defn

Let $G_1, G_2, H$ be groups. Consider maps $f_i:G_i\to H$. We want a group $G_1\star G_2$ such that there is a unique $f$ making the following diagram commute:
\[
\begin{tikzcd}
& G_1\star G_2\ar[d, dotted, "\exists!f"] & \\
G_1\ar[ur, "\iota_1"] \ar[r] & H & \ar[l] G_2\ar[ul, "\iota_2"'] \\
\end{tikzcd}
\]
Indeed, this is the coproduct in $\Grp$, and is the \underline{free product of the groups $G_1$ and $G_2$}. This can be given explicitly by using group presentations of $G_1, G_2$. If $G_1, G_2$ have presentations $G_i \cong \langle K_i \mid R_i \rangle$, where $K_i$ is the set of generators and $R_i$ the relations, then $G_1\star G_2$ has presentation $\langle K_1 \coprod K_2 \mid R_1 \coprod R_2 \rangle$.

\subsection*{\underline{At long last, SVK}}

What is $\pi_1(\mbb{S}^1\vee\mbb{S}^1)$? That is, what is the fundamental group of the figure 8? 

Last time, we showed that if $r$ is a retraction, then $\iota_*$ is injective and $r_*$ is surjective. So, because there is a retraction from $\mbb{S}^1\vee\mbb{S}^1$ to $\mbb{S}^1$, we know $\pi_1(\mbb{S}^1) ]\cong \Z$ injects into $\pi_1(\mbb{S}^1\vee\mbb{S}^1)$ somehow. 

Further, we get two different ways for $\Z$ to embed into it, because there are two different copies of $\mbb{S}^1$ to retract onto.

We're trying to make covering spaces of the figure 8. We eventually came to the Cayley graph for $\mbb{F}_2$. Because this is a tree, it has trivial fundamental group, so homotopy classes are determined by endpoints, and this clearly gives $\mbb{F}_2$.

$\mbb{F}_2$, the free group generated by $a$ and $b$, is $\pi_1(\mbb{S}^1\vee\mbb{S}^1)$. So $\pi_1(\mbb{S}^1\vee\mbb{S}^1)\cong\Z\star\Z$. Be careful: earlier we used $\star$ to refer to the join, which this is not.

\defn

A covering space $\begin{tikzcd} Y \ar[d, "p"] \\ X \end{tikzcd}$ such that $Y$ is simply connected is called the \underline{universal cover of $X$}. 


\thm 

If $\{(X_\alpha, x_\alpha)\}$ are ``nice," then 
\[
\pi_1(\bigvee_\alpha X_\alpha) = \star_\alpha\pi_1(X_\alpha)
\]
where again, $\star$ means free product. 

\proof

This is a direct corollary of Van Kampen's theorem. 

\section*{Lecture 17, 2/24/23} 

\thm (Siefert-Van Kampen)

\begin{enumerate}

\item If $X = \cup_{\alpha}A_\alpha$, with each $A_\alpha$ path connected and open, and such that there is some $x_0$ in each $A_\alpha$, and if, for all $\alpha, \beta, A_\alpha \cap A_\beta$ is path connected, then the natural map $\Phi:\star_\alpha\pi_1(A_\alpha, x_0)\to\pi_1(X, x_0)$

\item If, in addition, for all $\alpha,\beta,\gamma, A_\alpha \cap A_\beta \cap A_\gamma$ is path-connected, then the kernel of $\Phi$ is normally generated by elements of the form 
\[
i_{\alpha\beta}(\omega)i_{\beta\alpha}(\omega)^{-1}
\]
where $i_{\alpha\beta}:\pi_1(A_\alpha \cap A_\beta, x_0) \to \pi_1(A_\alpha, x_0)$.

\end{enumerate}

In other words, let $N$ be the subgroup normally generated by these elements (i.e. smallest normal subgroup containing these elements). Then $\pi_1(X, x_0) \cong \frac{\star_{\alpha}\pi_1(A_\alpha, x_0)}{N}$

\proof\,

\begin{enumerate}

\item Let $[f]\in\pi_1(X, x_0)$. Of course, $f:I\to X$ is a map from the interval to $X$ which is a loop based at $x_0$. We want to express this as the concatenation of finitely many loops, such that each loop is based at $x_0$, and stays entirely within $A_\alpha$, where the $\alpha$ depends on the curve. 

For every point in the image of $f$, take an open neighborhood. Without loss of generality this lies entirely in at least one of the $A_\alpha$. So we can pull this open set back by $f$. We can do this for every point, and obtain an open cover of $I$, which, by compactness, admits an finite subcover. This can be used to form finite cover of $I$, such that in each partition element $f$ does not leave one of the $A_\alpha$, and such that each open cover representing $A_\alpha$ overlaps with one representing $A_\beta$, and in this overlap, $f$ is in both. 

The intersections are where $f$ is in the intersection of the $A_\alpha$, and there we take wherever our curve is, and send it to $x_0$ and then send it back, making it a loop based at $x_0$. 

So we take a loop, decompose it into finitely segments, such that each of these segments lies entirely within an $A_\alpha$, and in the overlap between each segment, in the intersection, we draw a path to $x_0$ and then back (which is possible by path-connectedness), to make each segment a loop based at $x_0$, so the entire loop is their concatenation. 

Each concatenation is an element of the free product of the $\pi_1(A_\alpha, x_0)$, so the map $\Phi$ is onto. 

\item Suppose $[f]\in\star_\alpha\pi_1(A_\alpha)$, and suppose $\Phi([f]) = [c]$, where $c\equiv x_0$ is a constant map. We know that $[f]$ is the contatenation of loops, each of which stays entirely within one of the $A_\alpha$. Further, we know that $f$ extends to a map $f:\mbb{D}^2\to X$ because $[f] = [c]$. For every point in $I\times I$, it has an open neighborhood which lise entirely within the preimage of one of the $A_\alpha$, and so we can cover the disk/unit square in the same way we covered the unit interval before. 

So we have a finite open cover of $I\times I$. By real analysis, there is a Lebesgue number $\lambda$ such that any rectangle of side length $\lambda$ lives entirely in one of our open sets. We can cover $I\times I$ with rectangles of this size, and without loss of generality we can do this with at most triple intersections. 

We can cellulate this into a grid, such that each bottom edge of each grid cell is entirely within an overlap. We will pull a similar trick as before, in every vertex in $I \times I$, we run from that vertex to $x_0$, staying in the areas with triple intersection the entire time by staying in the edges.

So we can get a loop which is homotopic to $f$, by ``pushing it around" the squares. To ``push through" a square, we have to take a ``commutator" of the form $i_{\alpha\beta}(\omega)i_{\beta\alpha}(\omega)^{-1}$. 

\end{enumerate}

\qed

\rem

This is natural in the categorical sense, because the free product is the coproduct in group. Indeed, because the $A_\alpha$ include into $X$, we have a family of maps $(\iota_i)_*$, so by the universal property of the coproduct, there is a map $\Phi$ which makes the diagram below commute
\[
\begin{tikzcd}
& \star_\alpha\pi_1(A_\alpha)\ar[d, "\Phi"'] \\
\pi_1(A_\alpha)\ar[r, "(\iota_i)_*"] \ar[ur]& \pi_1(X)
\end{tikzcd}
\]

\section*{Lecture 18, 2/27/23}

Today we will see some applications of Van Kampen's theorem.  

\exm

We can think of the 2-sphere as the union of $U = \mbb{S}^2\setminus\text{north pole}$ and $V = \mbb{S}^2\setminus\text{south pole}$. We can see $U \cap V$ is homeomorphic to the cylinder, which has $\Z$ as its fundamental group. Call the generator $\omega$. Then $i_{\alpha\beta}(\omega) = 0$, and $i_{\beta\alpha}(\omega)^{-1} = 0$. 


Let $X$ be a space. If we glue on a disk (call the resulting space $Y$), we want to consider an open neighborhood of the image of the boundary of the disk under the attaching map. We want to consider $X$ plus the little bit of the disk as $A_\alpha$, and the interior of the disk as $A_\beta$. 

We can easily calculate $\pi_1(Y)$. 

By part (a) of Van Kampen's theorem, $\pi_1(A_\alpha,x_0) \star \pi_1(A_\beta, x_0)\to\pi_1(Y, x_0)$ is onto. But $\pi_1(A_\alpha, x_0) \cong \pi_1(X, x_0)$ because $A_\alpha$ deformation retracts onto $X$, and $\pi_1(A_\beta,x_0) = 0$. 

By part (b), $\pi_1(Y, x_0) = \frac{\pi_1(A_\alpha, x_0)}{N}$, where $N$ is normally generated by $\varphi(\del\mbb{D}^2)$, which is the inclusion of the generator of $\pi_1(A_\alpha \cap A_\beta)$ into $A_\alpha$ times the inverse of its inclusion into $A_\beta$. 

We have just proven the first part of the following proposition:

\prop\,

\begin{enumerate}

\item If $Y$ is obtained from $X$ by attaching 2-cells, then $\iota_*:\pi_1(X, x_0) \to \pi_1(Y, x_0)$ is onto. In fact, $\ker\iota_* = N$ and $N$ is generated by the conjugates of the loops which are the attaching maps. 

\item If $Y$ is obtained from $X$ by attaching $n$-cells, with $n > 2$, then $\iota_*:\pi_1(X)\to\pi_1(Y)$ is an isomorphism. 

\item If $X$ is a cell complex, $\pi_1(X) \cong \pi_1(X^2)$.

\end{enumerate}

\proof

\qed

Here is another application. 

Let $S$ be the shrinking wedge of circles, and consider the cone $C(S)$. Denote by ``the bad point" the point where all the circles are wedged together. 

Consider the disjoint union of three copies of $C(S)$, and add in a line from the bad point of each to an outside point. 

Let $A_\alpha$ be the first copy of $C(S)$, plus the line to the vertex, plus a little neighborhood of the vertex, and define $A_\beta, A_\gamma$ similarly. 

Note the intersection of any collection of these three is a small path-connected neighborhood of the vertex. 

So $\pi_1$ of this space is the free product of $C(S)$ with itself three times, which is trivial. 

This generalizes, and indeed if we connect any collection of spaces all to one point, then the fundamental group is the free product. 

\section*{Lecture 19, 3/1/23}

More applications of Van Kampen's theorem. 

\begin{enumerate}

\item Connected graphs: if we retract a spanning tree, we get a rose, which is just a vertex and some number of edges, i.e. a wedge of circles, so the fundamental group is the free product $\star_{i=1}^n\Z = F_n$. 

\item Closed, compact surface. Label the edges of a pentagon $aba^{-1}b^{-1}$, orienting them accordingly, then glue, we get a torus with a hole.
 

\rem

If you take a finite collection of polygons and orient and label their sides so that every label occurs exactly twice, and think of these labels and orientations as gluing instructions, then the natural quotient is a closed surface. 

\item Suppose we have a group $\langle a_1, b_1, a_2, b_2, \cdots, a_g, b_g \mid A \rangle$, where $A$ is one big word. Then we make a $4g$-gon, with edges labeled sequentially by $a_i,b_i,$ $a_i$ with the reverse orientation, then $b_i$ with the reverse orientation (we do this for orientability). So $A$ will be $[a_1,b_1][a_2,b_2]\cdots[a_g,b_g]$. 

We will wind up with lots of tori with holes, glued together, to get a genus $g$ surface. The group described earlier is thus $\pi_1$ of a genus $g$ surface, $M_g$.

\cor

If $g \neq h$, then $M_g$ is not homeomorphic to $M_h$

\proof

\qed

Similar methods as above give us fundamental group of $\RP^2, \pi_1(\RP^2)\cong \langle a \mid a^2 \rangle$. Similarly, the fundamental group of the Klein bottle is $\langle a, b \mid a^2b^2 \rangle$.

\cor

For every group $G$, there exists a 2-dimensional cell complex $X$ such that $\pi_1(X) = G$.

\proof

We construct it as follows. We start with a presentation $\langle G \mid R \rangle$, and begin with $|G|$-many loops. Then any word represents a specific loop in this space, so we cap off every word in $R$ with a disk, killing it.

\qed

\item We will think about the 3-sphere $\mbb{S}^3$ 


\end{enumerate}

\section*{Lecture 20, 3/4/23}

\section*{Lecture 21, 3/6/23}

\prop\,


\begin{enumerate}[label=(\alph*)]

\item If $p:(\tilde{X},\tilde{x_0})\to(X, x_0)$ is a covering map, then $p_*:\pi_1(\tilde{X},\tilde{x_0}) \to \pi_1(X, x_0)$ is injective.

\item The image of $p_*$ consist of equivalence classes of loops based at $x_0$ which lift to paths in $\tilde{X}$ starting at $\tilde{x_0}$, which remain loops.  

\end{enumerate}

\defn

Two covering spaces 
\[
\begin{tikzcd}
\tilde{X}_1\ar[dr, "p_1"] & & \tilde{X}_2\ar[dl, "p_2"'] \\
& X & \\
\end{tikzcd}
\]
are \underline{equivalent covering spaces} if there exist maps $f:\tilde{X}_1\to\tilde{X}_2, g:\tilde{X}_2\to\tilde{X}_1$, so that the diagram commutes:
\[
\begin{tikzcd}
\tilde{X}_1\ar[dr, "p_1"']\ar[rr, bend right = 20, "f"] & & \tilde{X}_2\ar[dl, "p_2"] \ar[ll, bend right = 15, "g"'']\\
& X & \\
\end{tikzcd}
\]

\defn

Two \underline{based} covering spaces are \underline{equivalent based covering spaces} if there are based maps $f, g$ that make the diagram commute:
\[
\begin{tikzcd}
(\tilde{X}_1,\tilde{x}_1)\ar[dr, "p_1"']\ar[rr, bend right = 20, "f"] & & (\tilde{X}_2,\tilde{x}_2)\ar[dl, "p_2"] \ar[ll, bend right = 10, "g"'']\\
& (X, x_0) & \\
\end{tikzcd}
\]

\defn 

An equality between covering spaces $(X, p), (X', p')$, is a \underline{symmetry} if $(X, p) = (X', p')$, and similarly for a symmetry of a based covering space. That is, a symmetry of a covering $p:X\to x$ is a map $f:X\to X$ such that $p \circ f = p$. 

\prop (Lifting criteria)
Given the following
\[
\begin{tikzcd}
& (\tilde{X}, \tilde{x}_0) \ar[d, "p"] \\
(Y, y) \ar[r, "f"] & (X, x) \\
\end{tikzcd}
\]
with $p$ a covering map, with $Y$ path connected and locally path connected, then $f_*(\pi_1(Y, y)) \subseteq p_*(\pi_1(\tilde{X}, \tilde{x}_0)$ if and only if there is a lift $\tilde{f}:(Y, y) \to (\tilde{X},\tilde{x}_0)$ that makes the diagram commute:
\[
\begin{tikzcd}
& (\tilde{X}, \tilde{x}_0) \ar[d, "p"] \\
(Y, y) \ar[r, "f"]\ar[ur, dotted, "\tilde{f}"] & (X, x) \\
\end{tikzcd}
\]

\proof\,

\subsection*{$=>$}

One direction is easy. Suppose that such a $\tilde{f}$ exists. Then $p\circ\tilde{f} = f$, so $(p\circ\tilde{f})_* = f_*$, so clearly the image of $f_*$ is a subset of the image of $p_*$.

\subsection*{$<=$} 

Now, suppose that $f_*(\pi_1(Y, y)) \subseteq p_*(\pi_1(\tilde{X},\tilde{x}_0))$. 

Let $\alpha:I\to Y$ be a path with endpoints $y, y'$. Then $f\circ\alpha$ is a path in $X$ with endpoints $x, x'$. By the homotopy lifting property, we can lift this to a path $\tilde{f\circ\alpha}$ with endpoints $\tilde{x},\tilde{x}'$, such that $f = p\circ\tilde{f\circ\alpha}$. 

Let $\alpha,\beta$ be two paths in $Y$ with the same endpoints. We lift this to two paths in $\tilde{X}$, and we want these paths to still have the same endpoints.

Consider the loop which we get from concatenating $\alpha$ and $\bar{\beta}$. The image of this loop under $f$ is a loop in $X$, and when we lift to $\tilde{X}$. This is independent of $\alpha$ because of the condition $f_*(\pi_1(Y, y)) \subseteq p_*(\pi_1(\tilde{X},\tilde{x}_0))$. 

Given a path $\alpha$ from $y$ to $y'$, we want to send $y'$ to the other endpoint of the lift of $\alpha$, and we have just shown this is well-defined (i.e. not dependent on choice of path $\alpha$, so this gives a well-defined function $\tilde{f}:Y\to \tilde{X}$. 

We now must show that this function is continuous. This comes from locally path connected: take an open set around $f(\alpha(1))$, and pull it back to an open set around $\alpha(1) = y'$. We can then lift without leaving the neighborhood that's evenly covered. 

\section*{Lecture 22, 3/8/23}

\prop

Given a covering $p:\tilde{X}\to X$ and a map $f:Y\to X$, with 2 lifts $\tilde{f}_1,\tilde{f}_2$ from $Y \to \tilde{X}$. If $\tilde{f}_1,\tilde{f}_2$ agree on one point $y\in Y,$ and $Y$ is connected, then $\tilde{f}_1,\tilde{f}_2$ agree on \underline{all} of $Y$.

\proof

Consider $\{y\in Y \mid \tilde{f}_1(y) = \tilde{f}_2(y)\}$. This will be open because of the covering space, and closed (this is a general fact, that the points of agreement of two functions is closed), so they must be either all of $Y$, or the empty set, and it is nonempty. So it is $Y$, so these are the same lift. 

\qed

Suppose we have a covering map $p:(\tilde{X},\tilde{x})\to(X, x)$. Let $H = \pi_1(\tilde{X},\tilde{x}), G = \pi_1(X, x)$. We have $H \subseteq G$, and indeed it is a subgroup. Let $g \in N_G(H)$, the normalizer of $H$, but not in $H$, and let $\tilde{g}$ be a lift of a loop representing it. Then this will lift to a path from $\tilde{x}$ to some other point $\tilde{x}'$. Indeed, there is a covering isomorphism $(\tilde{X},\tilde{x}) \to (\tilde{X},\tilde{x}')$. 

\subsection*{\underline{Creating $\tilde{X}$, the ``universal cover"}}

\rem

If $X$ has a simply connected cover $\tilde{X}$, then, for all $x \in X$, there exists an open set $U \ni x$ such that the map induced by the inclusion $\iota:U\hookrightarrow X$, $\pi_1(U)\hookrightarrow\pi_1(X)$ has trivial image.

Indeed, pick an evenly covered neighborhood $U_x$. This lifts to $\tilde{U}_{\tilde{x}}$. We investigate $\pi_1(\tilde{U})\to\pi_1(\tilde{X})$. By the definition of evenly covered, $\pi_1(\tilde{U}) \cong \pi_1(U)$, and the image of $\pi_1(\tilde{U})$ is trivial, as $\pi_1(\tilde{X})$ is trivial. So $p_*\circ\iota_*:\pi_1(U)\to\pi_1(X)$ has trivial image. This is the same map as the inclusion $U\hookrightarrow X$. 

\defn

If $X$ has the property that every point $x \in X$ admits such a neighborhood $U$ as above, then $X$ is said to be \underline{semi-locally simply connected}. 

\rem

Note that semilocally simply-connected is implied by locally simply connected, which is implied by locally contractible, which is implied by $X$ is a cell complex. So every (connected) cell complex has a universal cover. The shrinking wedge of circles is a good example of a space which is not semilocally simply-connected. 

\rem

Consider $\R^2\setminus\Q^2$, which, while still path connected, has a crazy fundamental group. This is very not semilocally simply-connected so (perhaps unsurprisingly) this does not have a universal cover. 

\rem

If a simply connected $\tilde{X}$ exists, then there exist bijections between the points in $\tilde{X}$ and $\{[\tilde{\gamma}] \mid \tilde{\gamma}$ is a homotopy equivalence class of a path in $\tilde{X}$ starting at $\tilde{x}\}$. This is true of all simply connected spaces. But this is the same as $\{[\gamma]\mid \gamma$ is a path in $X$ starting at $x\}$, from unique path lifting. 

\defn
\[
\tilde{X}\eqdef\{[\gamma] \mid \gamma \text{ is a path in $X$ starting at $x$}\}
\]

Is there a map $p:\tilde{X}\to X$? Consider $p:[\gamma]\mapsto \gamma(1)$. This is well defined by previous remarks. 

Let $\ms{U} = \{U \subseteq X \mid U$ is open, path connected, and $\pi_1(U)\to\pi_1(X)$ is trivial$\}$.

If $U \in\ms{U}$ and $V \subset U$ such that $V$ is open and path connected, then $V \in \ms{U}$.		

Note that if $X$ is path connected, locally path connected, and semilocally simply connected, $\ms{U}$ is a basis for the topology of $X$. 

\defn

Let $U_{[\gamma]} \subseteq \tilde{X}$ be $U_{[\gamma]} = \{[\gamma\star\eta]\mid \eta$ is a path in $U\subseteq X$ and $\eta(0) = \gamma(1)\}$. 

\section*{Lecture 23, 3/10/23}

Let $X$ be a path-connected/locally path-connected/s.l.s.c space. 

$\tilde{X} = \{[\gamma]\mid \gamma$ is a path in $X$ starting at $x_0\}$. 

$W = \{U\subseteq X \mid U$ is open, p.c., $\pi_1(U)\to\pi_1(X)$ is trivial $\}$. 

$W_{[\gamma]} = \{[\gamma\eta] \mid \eta$ is a path in $U$ with $\eta(0) = x_0\} \subseteq \tilde{X}$.

\rem 

If $X$ is p.c./l.p.c/s.l.s.c, then $W$ is a basis for the topology of $X$. 

\claim

The sets $\{U_{[\gamma]} \mid U \in W, \gamma$ any path with $\gamma(1) \in U \}$ forms a basis for a topology on $X$. The projection map $p:\tilde{X}\to X$, where $p([\gamma]) = \gamma(1)$, is a covering and $p|_{W_{[\gamma]}}:W_{[\gamma]}\to W$ is a homeomorphism. $p^{-1}(U) = \coprod U_{[\gamma]}$. Finally, $\pi_1(\tilde{X}, [x_0]) = \{0\}$. 

\proof

Some pictures?

\qed ?

\thm If $X$ is a pc/lpc/slsc space, then there exists a bijection between based coverse $p(\tilde{X},\tilde{x})\to(X,x)$ up to based isomorphism and subgroups of $\pi_1(X, x_0)$. 
\[
\left(
\begin{tikzcd}
(\tilde{X},\tilde{x_0}) \ar[dr]& \\
& (\tilde{X}, \tilde{x_0}) \\
(X, x_0)\ar[ur] & \\
\end{tikzcd}
\right)
\mapsto
\left(
\begin{tikzcd}
\{e\}\ar[dr, hook] & \\
& H \ar[dl, hook]\\
G = \pi_1(X, x_0) & \\
\end{tikzcd}
\right)
\]

\section*{Lecture 24, 3/15/23}

\exm

Find all covers of $\RP^2\vee\RP^2$

We know that there exists a two-sheeted cover
\[
\begin{tikzcd}
S^2 \ar[d, "p"] \\
\RP^2 \\
\end{tikzcd}
\]
We know further that $\pi_1(\RP^2\vee\RP^2) \cong \Z_2\star\Z_2 \cong \langle a, b \mid a^2 = b^2 = 1\rangle$. 

We know by the classification theorem that based covers (up to based isomorphism) are in bijection with the subgroups of $\Z_2\star\Z_2$. However, If we only want covers (up to (non-based!) isomorphism), then they are in bijection with the conjugacy classes of subgroups. 

The universal cover of $\RP^2\vee\RP^2$ is just a line of spheres glued together. 

Each sphere has two points corresponding to the wedge point of the two copies, so there are only so many ways to cover it. We can have a ray, a line, a circle, or a string.

There are two ways to make a ray (starting either with blue or red $\RP^2$), one way to make a line. There is then a circle for every even number $n$, because you need an even number of spheres. The final configuration is a string. This will consist of starting with either a red or blue $\RP^2$, and then a string of spheres. It can then end with either a blue or red $\RP^2$. So the string can be classified by the color of the ends, and how many spheres are in it. Note this includes the base case, with zero spheres. 

Jon claims that this is a complete list.

First, the universal cover is a normal cover. The deck transformation group is $\Z_2\star\Z_2$. 

Let's consider fundamental groups. For the circle configuration, $\pi_1$ will be $\Z$. For the ray, it's $\Z_2$. Of course, which copy of $\Z_2$ this projects onto depends on the choice of basepoints. This will project to the $\Z_2$ generated by a conjugate of $a$ or $b$, depending on red or blue. 

Normality: The line is normal, the rays are not (because there are distinguishable vertices). The circles are normal, and their deck transofmration groups correspond to finite dihedral groups. 

For the strings, the deck transformation group will be $\Z_2$ if the endpoints are the same color, and it won't be normal (unless there's only one sphere in between, in which case it'll be a 2-sheeted cover, meaning it can't help but be normal, because a subgroup having index 2 means it's normal). 

We now consider if the endpoints are different colors. Then, the deck transformation group will be trivial. This will not be normal. 























\end{document}